
\noindent\textbf{Related Works:} Although there have been significant research~\cite{resiRCA,chinchilla,Origin,footprint,intNAS, stateful, more, intermittentLearning} on enabling machine learning in intermittently powered devices, majority of them focus on performing inferences. Only intermittent learning~\cite{intermittentLearning} focuses on doing some training on the device, but with very small workloads and models. Considering the scale, scope and workload of our problem, we could not draw an equivalent comparison with those works, except for comparing their exemplar selection method (refer Figure~\ref{Fig:teachers}). Similarly, Ekya~\cite{ekya} only focuses on co-location of computation, and it's efficiency on finishing compute even on a custom hardware is shown in Figure~\ref{Fig:DVFSPrim}. 

\noindent\textbf{Alternate Power Sources:} A limitation of our work comes from the choice of solar energy: unavailability during night and bad weather makes the deployment harder. However, there has been significant recent  development in portable wind turbines~\cite{aero}, which can be deployed on rooftops, can work with $\ge 5mph$ wind speed, and can provide power equivalent of 15 solar cells. Therefore, similar technologies can be used to augment the harvesting mechanism. 

\noindent\textbf{Green Data Centers:} As sustainability has been getting significant traction, industrial partners have been contributing towards building green data centers~\cite{MSGreen,metaGreen}. Although using these data centers for computation can be an alternative, it will not solve the bandwidth and the privacy issues mentioned in \S\ref{sec:introduction}. Moreover, communicating and storing such high volume data will also require energy. Our solution decentralizes this massive compute using a sustainable approach and hence has its own merits. Further, this can help us build future solutions of using these decentralised nodes for other applications. We do encourage use of green data centers for other centralized compute applications.  

% \begin{itemize}
%     \item add about green data centers
%     \item add about alternative wind energy
%     \item add about generalization
%     \item add about related works
%     \item bring the 4 points (non-supervision, functionality, sustainability, and intermittency) together
% \end{itemize}